---
title: "EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### EDA

## Data source
https://www.kaggle.com/shivamb/netflix-shows
## Data in 
```{r, eval = FALSE}
unzip("netflix-shows.zip")
```

```{r}
library(data.table)
library(ggplot2)
data_in <- data.table(read.csv("netflix_titles_nov_2019.csv", 
                               stringsAsFactors = FALSE))

```

- Ratings

```{r}
ggplot(data_in, aes(x = rating)) +
  geom_histogram(stat="count") +
  ggtitle("Histogram of ratings")
# todo: Need to have a look at the different types of ratings
```
```{r}
library(dplyr)
library(tidyr)

countries <- data_in %>% 
  mutate(symbols = strsplit(country, ",")) %>% 
  unnest(symbols) %>%
  data.table()

```

```{r}

ggplot 
titles_by_countries 
```

- maps by country
- most popular actors as defined by how many titles they are in
- average number of actors quoted per title
- number of reviews per country
- do we have only one id per row?
- do we have duplicate names? 
- common bigrams / trigrams
- inverse frequency on the description
- count of characters in the description
- most popular groups
- most niche groups - the shows in them are not listed in many more groups on 
average

Zipf's Law ------------------------
Zipf’s law states that the frequency that a word appears is inversely 
proportional to its rank.
