---
title: "EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### EDA

## Data source
https://www.kaggle.com/shivamb/netflix-shows
## Data in 
```{r, eval = FALSE}
unzip("netflix-shows.zip")
```

```{r}
library(data.table)
data_in <- data.table(read.csv("netflix_titles_nov_2019.csv"))

```
```{r}
data_in[]
```

- Ratings by a country
```{r}
data_in[, mean(as.numeric(rating)), country]
```
- maps by country
- most popular actors as defined by how many titles they are in
- average number of actors quoted per title
- number of reviews per country
- do we have only one id per row?
- do we have duplicate names? 
- common bigrams / trigrams
- inverse frequency on the description
- count of characters in the description
- most popular groups
- most niche groups - the shows in them are not listed in many more groups on 
average

Zipf's Law ------------------------
Zipf’s law states that the frequency that a word appears is inversely 
proportional to its rank.
