---
title: "EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### EDA

## Data source
https://www.kaggle.com/shivamb/netflix-shows
## Data in 
```{r, eval = FALSE}
unzip("netflix-shows.zip")
```

```{r}
library(data.table)
library(ggplot2)
data_in <- data.table(read.csv("netflix_titles.csv", 
                               stringsAsFactors = FALSE))

# most of the obs of a data scientist - dealing with inconsistent / dirty 
# data
data_in[, date_added := gsub(", ", ",", date_added)]
data_in[, date_added := gsub("^ ", "", date_added)]
data_in[date_added == "", 
        date_added := paste0("January 01,", release_year)]

data_in[, date_added := as.Date(date_added, format = "%B %d,%Y")]
```

- Ratings

```{r}
ggplot(data_in, aes(x = rating)) +
  geom_histogram(stat = "count") +
  ggtitle("Histogram of ratings")
# todo: Need to have a look at the different types of ratings
```
```{r}
library(dplyr)
library(tidyr)

countries <- data_in %>% 
  mutate(symbols = strsplit(country, " ,")) %>%
  unnest(symbols) %>%
  mutate(symbols = strsplit(symbols, ",")) %>%
  unnest(symbols) %>%
  data.table()

countries[, symbols := gsub("^ ", "", symbols),]

```

```{r}

titles_by_countries <- countries[, .(title, symbols)]

counted_titles_per_country <- titles_by_countries[, .(counts = .N), .(symbols)]
setorder(counted_titles_per_country,-counts)

View(counted_titles_per_country)

counted_titles_per_country %>%
  top_n(15) %>%
  ggplot(aes(x = symbols, y = counts)) +
  geom_point()

```

Quite a lot of movies are produced / created in the United States. That makes 
sense from the point of view that the Western / Hollywood culture has been 
overtaking everything. On the other hand, we are talking a three times more
movies produces in the States than in the second most popular country - India.
It would be interesting to compare this to the countries of the viewers of 
Netflix.

- most popular actors as defined by how many titles they are in
```{r}
head(data_in)

actors <- data_in %>% 
  mutate(cast = strsplit(cast, " ,")) %>%
  unnest(cast) %>%
  mutate(cast = strsplit(cast, ",")) %>%
  unnest(cast) %>%
  data.table()

actors[, cast := gsub("^ ", "", cast),]

titles_by_actors <- actors[, .(title, cast)]

counted_titles_per_actor <- titles_by_actors[, .(counts = .N), .(cast)]
setorder(counted_titles_per_actor,-counts)

View(counted_titles_per_actor)

counted_titles_per_actor %>%
  top_n(10) %>%
  ggplot(aes(x = cast, y = counts)) +
  geom_point()

```
Lots of what I assume are Indian names.


- is there a good last name for an actor?
```{r}
library(tidytext)
counted_titles_per_actor[, id := 1:.N]
tokenized <- counted_titles_per_actor %>%
  unnest_tokens(word, cast)

tokenized %>%
  group_by(id) %>%
  count(id)

counted_titles_per_actor[, .(last_name = grep("", cast, value = TRUE))]
```


- most popular directors as defined by how many things they have directed + 
how many other directors were in it

```{r}
directors <- data_in %>% 
  mutate(director = strsplit(director, ", ")) %>%
  unnest(director) %>%
  mutate(director = strsplit(director, ",")) %>%
  unnest(director) %>%
  data.table()

# directors[, directors := gsub("^ ", "", directors),]

titles_by_directors <- directors[, .(title, director)]

counted_titles_per_directors <- titles_by_directors[, .(counts = .N), .(director)]
setorder(counted_titles_per_directors,-counts)

# View(counted_titles_per_actor)

counted_titles_per_directors %>%
  top_n(10) %>%
  ggplot(aes(x = director, y = counts)) +
  geom_point()

```

Curious - that is a lot more productions per director than I would have 
expected. Let's see how many of these were co-chaired.

```{r}
count_directors_per_title <- titles_by_directors[, .N, .(title)]
ggplot(count_directors_per_title, aes(x = N)) +
  geom_histogram()
```

There is a movie with 10 directors? That is a lot of people. Let see some more 
details.

```{r}
movie_titles <- count_directors_per_title[N > 4, title]

View(data_in[title %in% movie_titles])

```

Alright for the shows, if a bit surprising for the movies. Important to mention 
that there are prductions in this data set that have no directors. To be precise,
there are `r round(nrow(data_in[director == ""]) / nrow(data_in) * 100, 2)`%
of productions like that. 
Missing values can be sometimes imputed, but in this case there is nothing we
can really do about this lack besides be aware of it and try to take it into 
account. 


- maps by country
- average number of actors quoted per title
- number of reviews per country
- do we have only one id per row?
- do we have duplicate names? 
- common bigrams / trigrams
- inverse frequency on the description
- count of characters in the description
- most popular groups
- most niche groups - the shows in them are not listed in many more groups on 
average
- how does the distribution of movies change over time - movies produced in 
different countries added in 

- TV shows against movies over time
```{r}
# something sketchy happening here
counts_type <- data_in[, .N, .(type, date_added)]
max(counts_type$N)
counts_type[, quarter := zoo::as.yearqtr(date_added)]
counts_type <- counts_type[, .(N = sum(N)), .(type, quarter)]
max(counts_type$N)
ggplot(counts_type, aes(x = quarter, y = N, col = type)) +
  geom_line() +
  ggtitle("Additions of shows and movies by quarter")
```

Zipf's Law ------------------------
Zipf’s law states that the frequency that a word appears is inversely 
proportional to its rank.
