---
title: "EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### EDA

## Data source
https://www.kaggle.com/shivamb/netflix-shows
## Data in 
```{r, eval = FALSE}
unzip("netflix-shows.zip")
```

```{r}
library(data.table)
library(ggplot2)
data_in <- data.table(read.csv("netflix_titles.csv", 
                               stringsAsFactors = FALSE))

```

- Ratings

```{r}
ggplot(data_in, aes(x = rating)) +
  geom_histogram(stat = "count") +
  ggtitle("Histogram of ratings")
# todo: Need to have a look at the different types of ratings
```
```{r}
library(dplyr)
library(tidyr)

countries <- data_in %>% 
  mutate(symbols = strsplit(country, " ,")) %>%
  unnest(symbols) %>%
  mutate(symbols = strsplit(symbols, ",")) %>%
  unnest(symbols) %>%
  data.table()

countries[, symbols := gsub("^ ", "", symbols),]

```

```{r}

titles_by_countries <- countries[, .(title, symbols)]

counted_titles_per_country <- titles_by_countries[, .(counts = .N), .(symbols)]
setorder(counted_titles_per_country,-counts)

View(counted_titles_per_country)

counted_titles_per_country %>%
  top_n(15) %>%
  ggplot(aes(x = symbols, y = counts)) +
  geom_point()

```

Quite a lot of movies are produced / created in the United States. That makes 
sense from the point of view that the Western / Hollywood culture has been 
overtaking everything. On the other hand, we are talking a three times more
movies produces in the States than in the second most popular country - India.
It would be interesting to compare this to the countries of the viewers of 
Netflix.

- most popular actors as defined by how many titles they are in
```{r}
head(data_in)

actors <- data_in %>% 
  mutate(cast = strsplit(cast, " ,")) %>%
  unnest(cast) %>%
  mutate(cast = strsplit(cast, ",")) %>%
  unnest(cast) %>%
  data.table()

actors[, cast := gsub("^ ", "", cast),]

titles_by_actors <- actors[, .(title, cast)]

counted_titles_per_actor <- titles_by_actors[, .(counts = .N), .(cast)]
setorder(counted_titles_per_actor,-counts)

View(counted_titles_per_actor)

counted_titles_per_actor %>%
  top_n(10) %>%
  ggplot(aes(x = cast, y = counts)) +
  geom_point()

```
Lots of what I assume are Indian names. Let's make a cross of what actors
play in which countries.

This next code is not perfect -> for titles that have been produced in more than 
one country you will get duplicates in the actor counts. However, it does give 
us an idea where actors travel. 

```{r}

titles_by_countries_and_actors <- merge(
  titles_by_countries, titles_by_actors, 
  by = "title", allow.cartesian = TRUE)

tabled <- table(titles_by_countries_and_actors$cast,
                titles_by_countries_and_actors$symbols)

tabled_dt <- data.table(as.data.frame(tabled))

setnames(tabled_dt, c("Var1", "Var2"), c("actor", "country"))

tabled_dt <- merge(tabled_dt, counted_titles_per_actor, 
                   by.x = "actor", by.y = "cast")

tabled_dt <- tabled_dt[Freq > 0]

setorder(tabled_dt, -counts)

tabled_dt %>%
  filter(counts > 25) %>%
  # top_n(30) %>%
  ggplot(aes(x = actor, y = counts)) +
  facet_wrap(~country, scales = "free") +
  geom_point()

```


- is there a good last name for an actor?
```{r}
library(tidytext)
counted_titles_per_actor[, id := 1:.N]
tokenized <- counted_titles_per_actor %>%
  unnest_tokens(word, cast)

tokenized %>%
  group_by(id) %>%
  count(id) %>%
  
tokenized[, number_o]

counted_titles_per_actor[, .(last_name = grep("", cast, value = TRUE))]
```


- most popular directors as defined by how many things they have directed + 
how many other directors were in it


- maps by country
- average number of actors quoted per title
- number of reviews per country
- do we have only one id per row?
- do we have duplicate names? 
- common bigrams / trigrams
- inverse frequency on the description
- count of characters in the description
- most popular groups
- most niche groups - the shows in them are not listed in many more groups on 
average
- how does the distribution of movies change over time - movies produced in 
different countries added in 

Zipf's Law ------------------------
Zipf’s law states that the frequency that a word appears is inversely 
proportional to its rank.
